---
title: "HW 2: Visualization, Base R"
subtitle: "Data 180, Professor Bilen"
author: 
  name: Spencer Davis
email: davissp@dickinson.edu
date:  9/24/23
output: 
  html_document
---

Make sure that you have accepted the hw invitation link, and working with your clone repository.

**Due date:** September 26 beginning of class.

**Reminder:** You can insert R code in your Rmd file as a chunk in the following way:
```{r}
# Your code goes here
```
Make sure to insert a chunk for each (sub)section of the assignment.

**Theme of the assignment**: Let's practice visualization using base R. Load in the `Cars93` dataset by running the following code chunk:

```{r}
library(MASS)
data(Cars93)
```

The library function will load and attach add-on packages that have been previously installed on your machine. This will give you access to the functions and datasets available in the package. In general, you only need to install a package once (although it may require updating from time to time) but you need to access it using the library command every time you start a new session in R (and want the functionality). 

The questions below use the `Cars93` data frame as a source of data. This data frame is part of the MASS package and becomes available after loading it. You can get a view of the top of the data frame by using the head command. For more information, type `?Cars93`. Let's load `Cars93` in,
```{r}
head(Cars93)
```

# Base R
## Task 1
a) Create a Frequency table and Relative Frequency table (latter uses percentages) for the variable `Type`. Make sure to round all digits to 3 digits.
```{r}
table(Cars93$Type)
round (table(Cars93$Type)/nrow(Cars93), 3)

```
b) Create a frequency bar graph for the variable `Type`. You can have count on the y axis.
```{r}
barplot(table(Cars93$Type),ylab='count',xlab='Type',col = c( 'red','blue','yellow','green','purple','skyblue'), cex.lab=1.5, cex.names=1.2)
```

*	Your axes should be labeled and legible yes
*	The bars should be of different colors yes
*	The plot should include a title using the main argument yes
*	The names, labels, and axes sizes (cex parameter) should be set to 1.2 yes

c) Create a pie chart of the variable `Type`.
```{r}
pie(table(Cars93$Type),ylab='count',xlab='type of car', cex=1.2)
```

* The plot should include a title using the main argument.
*	The names, labels, and axes sizes (cex parameter) should be set to 1.2

d) What is the mode of the variable `Type`? Comment.
 "The mode is Midsize"

## Task 2 
a) Create a histogram of the variable `Horsepower`.
```{r}
hist(Cars93$Horsepower, col="deepskyblue",cex=1.2, main="Horsepower")
```

*	Your axes should be labeled
*	The hist body should have a color (your choice, not too dark to obscure lines) 
*	The plot should include a title using the main argument
*	The names, labels, and axes sizes should be set to 1.2

b) Create a second histogram with breaks at 25 HP increments instead of 50.
```{r}
hist(Cars93$Horsepower, col="deepskyblue",cex=1.2, main="Horsepower", breaks = 25,)
```

* Use the breaks argument to define the classes
* Other setting should be as in part (a)

c) Which histogram more closely follows or rule of thumb for the suggested number of classes? Explain.
```{r}
"I think the 2nd one fits better because we have a relalitivly large amount of data here."
```


## Task 3 
a) What is the type of variable `Passengers`? (Categorical nominal/ordinal or Numerical discrete/continuous?) What happens if we were to make a histogram for `Passengers`? What is the problem with this graph?
```{r}
typeof(Cars93$Passengers)
"It is Numerical Discrete, the problem is the outliers make the histogram less informitive."
```

b) Correct the problem in part (a) by creating a more appropriate visualization. Make sure to label your axes, place a title, choose a custom color as in the previous graphs. (Bonus: use factor() with levels option to display x axis values with zero observations on the plot. Hint: when you have discrete values, it might make sense to treat each value as a category.)
```{r}
barplot(table(Cars93$Passengers),ylab='count',xlab='Amount of Passengers',col = c( 'red','blue','yellow','green','purple','skyblue'), cex.lab=1.5)
```

## Task 4 
a) Make a histogram of the variable `Length` (the default breaks are fine). Other aspects of the plot should be as in part (a) in Task 2.
```{r}
hist(Cars93$Length, col="deepskyblue",cex=1.2, main="Length")
mean(Cars93$Length)
median(Cars93$Length)
```
```

b) What is the shape of the distribution?
"A wave shape starting small and getting bigger and ending small."

c) Find the mean and median of the variable `Length.` Do these answers meet your expectations from part (b)?
 "These match my ansers in part B because the average and the middle are almost the same. these shows how it is a wave with it starting small, middle big, and ending small again. "

You are done! 🏁 Don't forget to commit and push your .Rmd file to your Github repository before the due date.


